---
title: "R Notebook"
output: html_notebook
---

```{r}
suppressPackageStartupMessages(library("ggplot2"))
theme_set(ggpubr::theme_pubr(base_size=10, legend='bottom'))
```

```{r}
suppressPackageStartupMessages(library("DESeq2"))
```

# TCGA data
## Load raw expression data
```{r}
rawExpressionData = read.table("~/Downloads/TCGA_RNASeqRawCounts.txt", sep = " ", header = TRUE, quote = "")
```

```{r}
rawExpressionDataSubset = rawExpressionData[-c(60484:60488), 2:ncol(rawExpressionData)]
rownames(rawExpressionDataSubset) = gsub("\"", "", rawExpressionData$X..[-c(60484:60488)])
```

```{r}
rownames(rawExpressionDataSubset) = gsub("[[:punct:]].*","", rownames(rawExpressionDataSubset))
```

## Load the clinical data
```{r}
clinicalDataSampleId = read.table("~/Downloads/Clinical_details.xlsx - Sheet2.tsv", sep = "\t", header = TRUE, quote = "")
```

```{r}
#Make sure that the clinical data is in the same order as expression data
rownames(clinicalDataSampleId) = paste("X.", make.names(clinicalDataSampleId$sampleID), ".", sep = "")
```

```{r}
#clinical data with mutation information
clinicalData_mutation <- read.table("~/Downloads/skcm_tcga_pub_2015_clinical_data.tsv", sep = "\t", header = TRUE)
#clinicalData_BRAF<-clinicalData_BRAF[clinicalData_BRAF$Study.ID=='skcm_tcga',]
#Make sure that the clinical data is in the same order as expression data
rownames(clinicalData_mutation) <- paste("X.", make.names(clinicalData_mutation$Sample.ID), ".", sep = "")
```

Merge BRAF mutation information with clinical data
```{r}
clinicalDataSampleId$Mutation<-clinicalData_mutation[rownames(clinicalDataSampleId), "Mutation.Type"]
```

```{r}
clinicalDataSampleId$BRAF<- ifelse(clinicalDataSampleId$Mutation=="BRAF_Hotspot_Mutants", "Mutant", "WT")
clinicalDataSampleId$NRAS<- ifelse(clinicalDataSampleId$Mutation=="RAS_Hotspot_Mutants", "Mutant", "WT")
```

### Subset for tissue
```{r}
#tissue<-c("Regional Cutaneous or Subcutaneous Tissue (includes satellite and in-transit metastasis)", "Primary Tumor")
#tissue<-"Regional Lymph Node"
#rawExpressionDataSubsetTissue = rawExpressionDataSubset[, colnames(rawExpressionDataSubset) %in% rownames(clinicalDataSampleId)[clinicalDataSampleId$tumor_tissue_site %in% tissue]]

#clinicalData = clinicalDataSampleId[rownames(clinicalDataSampleId)%in%colnames(rawExpressionDataSubsetTissue), ]
#Make sure that the clinical data is in the same order as the sample expression data
seIdx = match(colnames(rawExpressionDataSubset), rownames(clinicalDataSampleId))
clinicalData = clinicalDataSampleId[seIdx, ]
```

## Perform Variance stablization transformation
### Prepare DESeq2 object
```{r}
#Could run it with full list of genes to make it comparable to the transformation applied on AVAST-M dataset
library("DESeq2")
dds <- DESeqDataSetFromMatrix(countData = rawExpressionDataSubset,
                              colData = clinicalData,
                              design = ~ 1)
dds
```
### Run vst 
```{r}
library("vsn")
#vsd = varianceStabilizingTransformation(dds, fitType="parametric")
#meanSdPlot(assay(vsd))
```
```{r}
#vsd = varianceStabilizingTransformation(dds, fitType="mean")
#meanSdPlot(assay(vsd))
```
```{r}
vsd = varianceStabilizingTransformation(dds, fitType="local")
meanSdPlot(assay(vsd))
```
Keeping the type local
```{r}
meanSdPlot(assay(vsd))
```

## Extract GRAMD1B information and calculate high/low groups based on mean value
```{r}
#vstNormalizedExpressionData = assay(vsd)
geneName<-"GRAMD1B"
select<-"ENSG00000023171"
#extract expression of this gene
expressionData <- data.frame(assay(vsd)[select, ])
#Replace ENSEMBL IDs with corresponding gene names
names(expressionData) <- geneName
stand.fun <- function(x){(x-mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE)}
stdSignature<-stand.fun(expressionData$GRAMD1B)
names(stdSignature)<-rownames(expressionData) #make sure they retain the name of the samples
clinicalData$Signature<- stdSignature[rownames(clinicalData)] #merge signature in clinical data frame
```

```{r}
#Divide in high/low groups based on mean value
co_mean<-mean(clinicalData$Signature)
clinicalData$SignatureGroupMean <- as.factor(ifelse(clinicalData$Signature >= as.numeric(co_mean), "High", "Low"))
```

## Repeat for PTEN information and calculate high/low groups based on mean value
```{r}
#vstNormalizedExpressionData = assay(vsd)
geneName<-"PTEN"
select<-"ENSG00000171862"
#extract expression of this gene
expressionData <- data.frame(assay(vsd)[select, ])
#Replace ENSEMBL IDs with corresponding gene names
names(expressionData) <- geneName
stand.fun <- function(x){(x-mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE)}
stdSignature<-stand.fun(expressionData$PTEN)
names(stdSignature)<-rownames(expressionData) #make sure they retain the name of the samples
clinicalData$PTEN<- stdSignature[rownames(clinicalData)] #merge signature in clinical data frame
```

```{r}
#Divide in high/low groups based on mean value
co_mean<-mean(clinicalData$PTEN)
clinicalData$PTENGroupMean <- as.factor(ifelse(clinicalData$PTEN >= as.numeric(co_mean), "High", "Low"))
```

## Check interaction of GRAMD1B with BRAF
```{r}
clinicalData$BRAF<-factor(clinicalData$BRAF)
#levels(clinicalData$BRAF)<-c("WT", "Mutant")
my_comparisons <- list( c("WT", "Mutant"))
ggplot(clinicalData[!is.na(clinicalData$BRAF),], aes(x=BRAF, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "GRAMD1B risk group")+
  xlab("BRAF")+
  ylab("VST normalized GRAMD1B\nexpression (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")
ggsave(paste("../results/TCGA_BRAF.pdf", sep=""), device = "pdf", width =8, units = "cm", height=6)
```
## Check interaction of GRAMD1B with NRAS
```{r}
clinicalData$NRAS<-factor(clinicalData$NRAS)
#levels(clinicalData$BRAF)<-c("WT", "Mutant")
my_comparisons <- list( c("WT", "Mutant"))
ggplot(clinicalData[!is.na(clinicalData$NRAS),], aes(x=NRAS, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "GRAMD1B risk group")+
  xlab("NRAS")+
  ylab("VST normalized GRAMD1B expression\n(stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")
ggsave(paste("../results/TCGA_NRAS.pdf", sep=""), device = "pdf", width =8, units = "cm", height=6)
```

```{r}
clinicalData$PTENGroupMean<-factor(clinicalData$PTENGroupMean)
#levels(clinicalData$BRAF)<-c("WT", "Mutant")
my_comparisons <- list( c("High", "Low"))
ggplot(clinicalData[!is.na(clinicalData$PTENGroupMean),], aes(x=PTENGroupMean, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "GRAMD1B risk group")+
  xlab("PTEN")+
  ylab("VST normalized GRAMD1B expression (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")
ggsave(paste("../results/TCGA_PTEN.pdf", sep=""), device = "pdf", width =8, units = "cm", height=6)
```
## BRAF mutation and PTEN loss
```{r}
clinicalData$BRAF_PTEN<-ifelse(
   (clinicalData$BRAF=="Mutant") & (clinicalData$PTENGroupMean=="Low"),
   "Yes",
   "No")
clinicalData$BRAF_PTEN<-factor(clinicalData$BRAF_PTEN)
my_comparisons <- list( c("Yes", "No"))
ggplot(clinicalData[!is.na(clinicalData$BRAF_PTEN),], aes(x=BRAF_PTEN, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "GRAMD1B risk group")+
  xlab("BRAF mutant + PTEN low")+
  ylab("VST normalized GRAMD1B expression (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")
ggsave(paste("../results/TCGA_BRAF_PTEN.pdf", sep=""), device = "pdf", width =8, units = "cm", height=6)
```
```{r}
clinicalData$BRAF_PTEN_NRAS<-ifelse(
   (clinicalData$BRAF=="Mutant") & (clinicalData$PTENGroupMean=="Low") & (clinicalData$NRAS==""),
   "Yes",
   "No")
clinicalData$BRAF_PTEN_NRAS<-factor(clinicalData$BRAF_PTEN_NRAS)
my_comparisons <- list( c("Yes", "No"))
ggplot(clinicalData[!is.na(clinicalData$BRAF_PTEN_NRAS),], aes(x=BRAF_PTEN_NRAS, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "GRAMD1B risk group")+
  xlab("BRAF mutant + PTEN low")+
  ylab("VST normalized GRAMD1B expression (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")
```

## Perform survival analysis
```{r}
library('survival')
clinicalData$survival_d_rc = Surv(time  = (clinicalData$X_OS)/365.25,
                              event = as.numeric(clinicalData$X_OS_IND))
```

### Univariate survival analysis
```{r}
library(survminer)
fit <- survfit( survival_d_rc ~ SignatureGroupMean, 
      data = clinicalData )
# KM curve for signature
pdf(paste("~/Desktop/Melanoma/Shruthy/GRAMD1B/results/survival/TCGA_all_OS_K-M_curve.pdf", sep=""), width = 3.15, height = 3.15, family="sans", pointsize=7)
plot(fit, col = c("blue", "red"), xlab = "Time (years)", ylab = "Survival probability", lwd=2)
pvalStr = survminer::surv_pvalue(fit = fit, data = clinicalData)$pval
if(pvalStr>=0.0001){
   pvalStr=round(pvalStr, digits = 4)
} else{
  pvalStr=formatC(pvalStr, format = "e", digits = 2) 
}
text(x= 5, y=0.1, paste0("p=", pvalStr))
legend("topright", col = c("blue", "red"), legend = c("High", "Low"), lty = 1, lwd=2)
dev.off()
```

### Multivariate survival analysis
```{r}
tcga_surv = data.frame()
fit = coef(summary(coxph(survival_d_rc ~ Signature,data=clinicalData)))
mid  = fit["Signature",c("exp(coef)")]
low  = exp(fit["Signature",c("coef")]-
              qnorm(.975)*fit["Signature",c("se(coef)")])                   
high = exp(fit["Signature",c("coef")]+
              qnorm(.975)*fit["Signature",c("se(coef)")])  
pval = fit["Signature",c("Pr(>|z|)")]
tcga_surv["Signature",c("HR","low","high","pval")] = c(mid,low,high,pval)

fit = coef(summary(coxph(survival_d_rc ~ Signature+gender+age_at_initial_pathologic_diagnosis+pathologic_stage,
                         data=clinicalData)))
mid  = fit["Signature",c("exp(coef)")]
low  = exp(fit["Signature",c("coef")]-
              qnorm(.975)*fit["Signature",c("se(coef)")])                   
high = exp(fit["Signature",c("coef")]+
              qnorm(.975)*fit["Signature",c("se(coef)")])  
pval = fit["Signature",c("Pr(>|z|)")]
tcga_surv["Signature+Sex+Age+Stage",c("HR","low","high","pval")] = c(mid,low,high,pval)

fit = coef(summary(coxph(survival_d_rc ~ Signature+gender+age_at_initial_pathologic_diagnosis+breslow_depth_value,
                         data=clinicalData)))
mid  = fit["Signature",c("exp(coef)")]
low  = exp(fit["Signature",c("coef")]-
              qnorm(.975)*fit["Signature",c("se(coef)")])                   
high = exp(fit["Signature",c("coef")]+
              qnorm(.975)*fit["Signature",c("se(coef)")])  
pval = fit["Signature",c("Pr(>|z|)")]
tcga_surv["Signature+Sex+Age+Breslow",c("HR","low","high","pval")] = c(mid,low,high,pval)

tcga_surv$Group = rownames(tcga_surv)
```

```{r}
tcga_surv$Group <- factor(tcga_surv$Group, levels = c("Signature+Sex+Age+Breslow", "Signature+Sex+Age+Stage", "Signature"))

if(pvalStr>=0.0001){
   pvalStr=round(pvalStr, digits = 4)
} else{
  pvalStr=formatC(pvalStr, format = "e", digits = 2) 
}

g1 <- ggplot(data=tcga_surv,
       aes(x = Group,y = HR, ymin = low, ymax = high, label = #paste("p-value=",formatC(pval,format="e"), sep = "")))+
   ifelse(pval>=0.0001, paste("p=",round(pval, 4), sep = ""), paste("p=",formatC(pval, format = "e", digits = 2), sep = ""))))+
   geom_pointrange()+
   geom_hline(yintercept =1, linetype=2)+
   xlab('')+
   ylab('')+
   ylab("Hazard Ratio (95% Confidence Interval)")+
   geom_errorbar(aes(ymin=low, ymax=high),width=0.2,cex=1)+ 
   ylim(c(0,4))+
   geom_text(hjust = 1.1, vjust=-1.25, size=2.5,  family="sans")+
   coord_flip()+
   theme(text=element_text(size=7,  family="sans"))
   #theme_set(ggpubr::theme_pubr(base_size=10, legend='bottom'))
ggsave(paste("~/Desktop/Melanoma/Shruthy/GRAMD1B/results/survival/TCGA_all_OS_ForestPlots.pdf", sep=""), device = "pdf", width = 10, units = "cm", height=4.5)
```

## Divide Signature based on Breslow thickness
```{r}
clinicalData$breslow_depth_value_group<-NA
clinicalData$breslow_depth_value_group[clinicalData$breslow_depth_value<=1]<-"<=1 mm"
clinicalData$breslow_depth_value_group[(clinicalData$breslow_depth_value>1)&
                                          (clinicalData$breslow_depth_value<=2)]<-">1-2 mm"
clinicalData$breslow_depth_value_group[(clinicalData$breslow_depth_value>2)&
                                          (clinicalData$breslow_depth_value<=3)]<-">2-3 mm"
clinicalData$breslow_depth_value_group[(clinicalData$breslow_depth_value>3)&
                                          (clinicalData$breslow_depth_value<=4)]<-">3-4 mm"
clinicalData$breslow_depth_value_group[(clinicalData$breslow_depth_value>4)]<-">4 mm"

```

```{r}
clinicalData$SignatureGroupMean_breslow<-NA
clinicalData$SignatureGroupMean_breslow[(clinicalData$SignatureGroupMean=="High")&                                        (clinicalData$breslow_depth_value_group%in%c(">3-4 mm", ">4 mm"))]<-"High GRAMD1B/Breslow>3mm"
clinicalData$SignatureGroupMean_breslow[(clinicalData$SignatureGroupMean=="Low")&                                        (clinicalData$breslow_depth_value_group%in%c(">3-4 mm", ">4 mm"))]<-"Low GRAMD1B/Breslow>3mm"
```

```{r}
fit <- survfit( survival_d_rc ~ SignatureGroupMean_breslow, 
      data = clinicalData )
# KM curve for signature
pdf(paste("~/Desktop/Melanoma/Shruthy/GRAMD1B/results/survival/TCGA_all_OS_K-M_curve_breslow.pdf", sep=""), width = 3.15, height = 3.15, family="sans", pointsize=7)
plot(fit, col = c("blue", "red"), xlab = "Time (years)", ylab = "Survival probability", lwd=2)
pvalStr = survminer::surv_pvalue(fit = fit, data = clinicalData)$pval
if(pvalStr>=0.0001){
   pvalStr=round(pvalStr, digits = 4)
} else{
  pvalStr=formatC(pvalStr, format = "e", digits = 2) 
}
text(x= 5, y=0.1, paste0("p=", pvalStr))
legend("topright", col = c("blue", "red"), legend = c("High", "Low"), lty = 1, lwd=2)
dev.off()
```