---
title: "R Notebook"
output: html_notebook
---

```{r}
suppressPackageStartupMessages(library("ggplot2"))
theme_set(ggpubr::theme_pubr(base_size=10, legend='bottom'))
```

```{r}
suppressPackageStartupMessages(library("DESeq2"))
```

#Note that we do not have Clark Scores or MIA scores for AVAST-M lymph node samples and these need to be commented out to avoid errors. 
```{r}
#Load the data
tissue<-c("Skin")
load(paste("~/Desktop/Melanoma/deseq2Results/tc",tissue,"EventMetNo_VS_tc",tissue,"EventMetYes_CovariateCorrection.deseq2/de.Rdata", sep=""))
```

```{r}
#find ENSEMBL ID corresponding to GRAMD1B
geneName<-"GRAMD1B"
select<-as.character(res.annot$id[res.annot$Name==geneName])
#extract expression of this gene
expressionData <- data.frame(assay(vsd)[select, ])
#Replace ENSEMBL IDs with corresponding gene names
names(expressionData) <- geneName
stand.fun <- function(x){(x-mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE)}
stdSignature<-stand.fun(expressionData$GRAMD1B)
names(stdSignature)<-rownames(expressionData) #make sure they retain the name of the samples
```

```{r}
#Multiply by beta coefficient?
#betaCoeff <- read.table("~/Desktop/Melanoma/betaCoeff.tsv", sep = "\t", header = TRUE, quote = "")
#extract beta coefficient for the gene of interest
#betaCoeffGene <- betaCoeff[select, "EventMet_Yes_vs_No"]
#weightedSignature <- expressionData*betaCoeffGene
#standardization of the weighted signature
#stand.fun <- function(x){(x-mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE)}
#stdWeightedSignature <- stand.fun(weightedSignature$GRAMD1B)
#names(stdWeightedSignature)<-rownames(weightedSignature) #make sure they retain the name of the samples
```

As expected the absolute values after standardization of weighted signature (stand.fun(weightedSignature\$GRAMD1B)) is same as those of standardized expression values (stand.fun(expressionData\$GRAMD1B)). The only difference comes from sign of the beta coefficient which is negative for GRAMD1B

```{r}
#Load clinical data
clinicalData<-data.frame(colData(vsd))
clinicalData$Signature<- stdSignature[rownames(clinicalData)] #merge signature in clinical data frame
```

#Check association of GRAMD1B with relapse
```{r}
my_comparisons <- list( c("Yes", "No"))
ggplot(clinicalData[!is.na(clinicalData$EventMet),], aes(x=EventMet, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("Relapse (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")#+ # Add pairwise comparisons p-value
  #ggpubr::stat_compare_means(label.y = 6, method = "anova", size = 2.5, family="sans")#+

```
Find suitable cut-off. After talking to Roy, it seems that we could go for a weighted GRAMD1B value of 0?
Let's check the median value. May be I can take the median cut-off and perform chi-squared test to check if there is no difference between the mean of 2 groups?
```{r}
#Divide in high/low groups based on mean and median
co_mean<-mean(clinicalData$Signature)
co_median<-median(clinicalData$Signature)
co_0_33<-quantile(clinicalData$Signature, 0.33)
co_0_66<-quantile(clinicalData$Signature, 0.66)
if (tissue=="Skin") {
  co_density_intersection<--0.223156179#0.235202939
}else{
  co_density_intersection<-0.3481057115#-0.223156179#0.235202939
}
suppressPackageStartupMessages(library(plotly))
p<-ggplot(data = clinicalData, aes(x = Signature, fill=EventMet)) +
  geom_density(alpha=0.3)+
  geom_vline(xintercept=c(co_mean, co_median, co_0_33, co_0_66, co_density_intersection), linetype=c('twodash', 'longdash', 'dotted', 'dotdash', 'dashed'))+
  ggtitle(tissue)
ggplotly(p) 
```

```{r}
clinicalData$SignatureGroupMean <- as.factor(ifelse(clinicalData$Signature >= as.numeric(co_mean), "High", "Low"))
clinicalData$SignatureGroupMedian <- as.factor(ifelse(clinicalData$Signature >= as.numeric(co_median), "High", "Low"))
clinicalData$SignatureGroup0_33 <- as.factor(ifelse(clinicalData$Signature >= as.numeric(co_0_33), "High", "Low"))
clinicalData$SignatureGroup0_66 <- as.factor(ifelse(clinicalData$Signature >= as.numeric(co_0_66), "High", "Low"))
clinicalData$SignatureGroupDensityIntersection <- as.factor(ifelse(clinicalData$Signature >= as.numeric(co_density_intersection), "High", "Low"))

print(chisq.test(clinicalData$SignatureGroupMean, clinicalData$EventMet))
print(chisq.test(clinicalData$SignatureGroupMedian, clinicalData$EventMet))
print(chisq.test(clinicalData$SignatureGroup0_33, clinicalData$EventMet))
print(chisq.test(clinicalData$SignatureGroup0_66, clinicalData$EventMet))
print(chisq.test(clinicalData$SignatureGroupDensityIntersection, clinicalData$EventMet))
```
#Check association of NRAS with other clinical covariates
```{r}
print(chisq.test(clinicalData$NRAS, clinicalData$EventMet))
print(chisq.test(clinicalData$NRAS, clinicalData$Bres))
print(chisq.test(clinicalData$NRAS, clinicalData$Ulc))
```
#Check association of GRAMD1B with ulceration
```{r}
my_comparisons <- list( c("Present", "Absent"))
ggplot(clinicalData[!is.na(clinicalData$Ulc),], aes(x=Ulc, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("Ulc (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")#+ # Add pairwise comparisons p-value
  #ggpubr::stat_compare_means(label.y = 6, method = "anova", size = 2.5, family="sans")#+
```

#Check association of GRAMD1B with Bres
```{r}
my_comparisons <- list( c("<= 2.0 mm", ">2-4mm"), c("<= 2.0 mm", ">4.0mm"), c(">2-4mm", ">4.0mm"))
ggplot(clinicalData[!is.na(clinicalData$Bres),], aes(x=Bres, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("Bres (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")+ # Add pairwise comparisons p-value
  ggpubr::stat_compare_means(label.y = 6, method = "anova", size = 2.5, family="sans")#+
```

#Check association of GRAMD1B with Stage
```{r}
my_comparisons <- list( c("IIB", "IIC"), c("IIB", "IIIA"), c("IIB", "IIIB"), c("IIB", "IIIC"),
                        c("IIC", "IIIA"), c("IIC", "IIIB"), c("IIC", "IIIC"),
                        c("IIIA", "IIIB"), c("IIIA", "IIIC"),
                        c("IIIB", "IIIC"))
ggplot(clinicalData[!is.na(clinicalData$Stage),], aes(x=Stage, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("Stage (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")+ # Add pairwise comparisons p-value
  ggpubr::stat_compare_means(label.y = 9, method = "anova", size = 2.5, family="sans")#+
```
```{r}
ggplot(clinicalData[!is.na(clinicalData$Stage_binary),], aes(x=Stage_binary, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("Stage (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(method = "t.test", size = 2.5, family="sans")#+ # Add pairwise comparisons p-value
  #ggpubr::stat_compare_means(label.y = 9, method = "anova", size = 2.5, family="sans")
```

#Check association of GRAMD1B with Site
```{r}
my_comparisons <- list( c("Head and neck", "Lower lims"), c("Head and neck", "Trunk"), c("Head and neck", "Upper limbs"),
                        c("Lower lims", "Trunk"), c("Lower lims", "Upper limbs"),
                        c("Trunk", "Upper limbs"))
ggplot(clinicalData[!is.na(clinicalData$Site),], aes(x=Site, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("Site (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")+ # Add pairwise comparisons p-value
  ggpubr::stat_compare_means(label.y = 9, method = "anova", size = 2.5, family="sans")#+

```

#Check association of GRAMD1B with BRAF mutation
```{r}
my_comparisons <- list( c("V600E", "WT"))
ggplot(clinicalData[!is.na(clinicalData$BRAF),], aes(x=BRAF, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("BRAF (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")#+ # Add pairwise comparisons p-value
  #ggpubr::stat_compare_means(label.y = 9, method = "anova", size = 2.5, family="sans")#+
```

#Check association of GRAMD1B with NRAS mutation
```{r}
my_comparisons <- list( c("Mutant", "WT"))
ggplot(clinicalData[!is.na(clinicalData$NRAS),], aes(x=NRAS, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("NRAS (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")#+ # Add pairwise comparisons p-value
  #ggpubr::stat_compare_means(label.y = 9, method = "anova", size = 2.5, family="sans")#+
```

#Check association of GRAMD1B with TIL counts
```{r}
jt_ns_sm_scores = read.table("../JT_NS_SM_Combined.xls - JT_NS_SM.tsv", sep = "\t", header = TRUE, quote = "")
combinedScore <- data.frame("Scanned_file.me"= jt_ns_sm_scores$JT_Scanned_file.me, 
                                "R.Seq_sampleID"= jt_ns_sm_scores$JT_R.Seq_sampleID, 
                                "ClarkScore"= ifelse(jt_ns_sm_scores$JT_Clark_score==jt_ns_sm_scores$NS_Clark_score,
                            jt_ns_sm_scores$JT_Clark_score, 
                            jt_ns_sm_scores$SM_Clark_score),
                           "MIAScore"= ifelse(jt_ns_sm_scores$JT_TIL_grade==jt_ns_sm_scores$NS_TIL.GRADE,
                            jt_ns_sm_scores$JT_TIL_grade, 
                            jt_ns_sm_scores$SM_TIL_grade))
#Remove duplicated entries
combinedScore<-combinedScore[!duplicated(combinedScore$R.Seq_sampleID), ]
rownames(combinedScore)<-combinedScore$R.Seq_sampleID #rename rownames with sample ids
```

```{r}
rownames(clinicalData)<-clinicalData$RNA.Seq.Sample
df<-data.frame("EventMet"=clinicalData$EventMet,
               "ClarkScore"=combinedScore[rownames(clinicalData), "ClarkScore"],
               "MIAScore"=combinedScore[rownames(clinicalData), "MIAScore"],
               "Signature"=clinicalData$Signature)
rownames(df)<-rownames(clinicalData)

df$ClarkScore <- as.factor(df$ClarkScore)
levels(df$ClarkScore) <- c("absent", "nonbrisk", "brisk")
```

```{r}
my_comparisons <- list( c("absent", "nonbrisk"), c("nonbrisk", "brisk"), c("absent", "brisk") )
#df$JT_Clark_score <- factor(df$JT_Clark_score, levels = c("absent", "nonbrisk", "brisk"))
ggplot(df[!is.na(df$ClarkScore), ], aes(x=ClarkScore, y=Signature#, color=DRBrain
                                        ))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("Clark score (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(df$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")+ # Add pairwise comparisons p-value
  ggpubr::stat_compare_means(label.y = 6, method = "anova", size = 2.5, family="sans")#+
 # facet_wrap(~DRBrain)# Add global p-value
```

```{r}
my_comparisons <- list( c("0", "1"), c("0", "2"), c("0", "3"), c("1", "2"), c("1", "3"), c("2", "3"))
df$MIAScore <- as.factor(df$MIAScore)
ggplot(df[!is.na(df$MIAScore), ], aes(x=MIAScore, y=Signature#, color=DRBrain
                                      ))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5)+
  scale_fill_brewer(type="qual", palette = "Dark2", name = "MIA score")+
  xlab("MIA score (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(df$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(comparisons = my_comparisons, method = "t.test", size = 2.5, family="sans")+ # Add pairwise comparisons p-value
  ggpubr::stat_compare_means(label.y = 6, method = "anova", size = 2.5, family="sans")#+
  #facet_wrap(~DRBrain)
     # Add global p-value

```
```{r}
#my_comparisons <- list( c("Mutant", "WT"))
ggplot(clinicalData[!is.na(clinicalData$treatment),], aes(x=treatment, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("Treatment (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(method = "t.test", size = 2.5, family="sans")#+ # Add pairwise comparisons p-value
  #ggpubr::stat_compare_means(label.y = 9, method = "anova", size = 2.5, family="sans")#+
```
```{r}
#my_comparisons <- list( c("Mutant", "WT"))
ggplot(clinicalData[!is.na(clinicalData$ECOG),], aes(x=ECOG, y=Signature))+
  geom_violin()+
  geom_jitter(height = 0, width = 0.1, alpha = 0.5, aes())+
  #scale_fill_brewer(type="qual", palette = "Dark2", name = "Cam_121 risk group")+
  xlab("ECOG (AVAST-M Skin)")+
  ylab("Weighted GRAMD1B score (stand.)")+
  #geom_hline(yintercept = quantile(clinicalData$Signature, 0.75), color = "grey", linetype = "longdash")+
  theme(text=element_text(size=7,  family="sans"))+
  ggpubr::stat_compare_means(method = "t.test", size = 2.5, family="sans")#+ # Add pairwise comparisons p-value
  #ggpubr::stat_compare_means(label.y = 9, method = "anova", size = 2.5, family="sans")#+
```

#Perform differential expression analysis
```{r}
combinedScore$ClarkScore <- as.factor(combinedScore$ClarkScore)
levels(combinedScore$ClarkScore) <- c("absent", "nonbrisk", "brisk")
clinicalData$ClarkScore<-combinedScore[rownames(clinicalData), "ClarkScore"]
#merge(clinicalData, combinedScore, by.x="RNA.Seq.Sample", by.y="R.Seq_sampleID")
#rownames(clinicalData)<-clinicalData$RNA.Seq.Sample
#clinicalData$ClarkScore=combinedScore$ClarkScore[rownames(clinicalData)]
```

```{r}
#While accounting for Stage, NRAS mutation, TIL score and EventMet
#Subset to only those samples which have non-missing entires for these covariates
clinicalData_sub<-clinicalData[(!is.na(clinicalData$Stage))&(!is.na(clinicalData$NRAS))&(!is.na(clinicalData$ClarkScore))&(!is.na(clinicalData$EventMet)), ]
data.f_sub<-data.f[, rownames(clinicalData_sub)]
```

```{r}
table(clinicalData_sub$SignatureGroupMean)
```

```{r}
dds1 <- DESeqDataSetFromMatrix(countData = data.f_sub,
                               colData   = clinicalData_sub,
                               design    = ~ Stage+NRAS+ClarkScore+EventMet+SignatureGroupMean)
```

```{r}
totalcounts1.gene = rowSums(counts(dds1))
dds1 <- dds1[totalcounts1.gene>=10,]
```

```{r}
# dimensions post filtering
m = nrow(dds1)
n = ncol(dds1)
```

```{r}
dds1 <- DESeq(dds1)
res1 <- results(dds1)
res1
```
```{r}
#Merge with gene names and check if the differentially expressed genes make sense
res<-data.frame(res1)
res<-res[order(res$padj), ]
res$Name<-res.annot[rownames(res), "Name"]
head(res)
#GRAMD1B is differentially expressed in GRAMD1B groups and is downregulated in the low expression group compared to the high-expression group so atleast this makes sense.
```
#Perform lfc shrinkage for GSEA
```{r}
library("apeglm")
coefName = resultsNames(dds1)
resWithLfcShrinkage = data.frame(lfcShrink(dds1, coef = tail(coefName, n=1), type="apeglm"))
#Add suffix to colnames for easy identification later on.
colnames(resWithLfcShrinkage) = paste(colnames(resWithLfcShrinkage),"lfcShrinkApplied", sep = "_")
#Add ENSEMBL ID as a column in the data frame to ease merging later
resWithLfcShrinkage$id = rownames(resWithLfcShrinkage)
```

```{r}
#Merge these with DE original DE results
res$id<-rownames(res)
deResultsUpdated = merge(x = res, y = data.frame(resWithLfcShrinkage), by = "id")
deResultsUpdated = deResultsUpdated[order(deResultsUpdated$padj),]
#Save the results
write.table(deResultsUpdated[deResultsUpdated$padj_lfcShrinkApplied<0.1, ], file = "../results/de_gsea/deWithLfcSkrinkageApplied_GRAMD1B_all.tsv", col.names = TRUE, row.names = FALSE, quote = FALSE, sep = '\t')
#Save the differentially expressed genes (padj<0.1)
write.table(deResultsUpdated[deResultsUpdated$padj_lfcShrinkApplied<0.1, ], file = "../results/de_gsea/deWithLfcSkrinkageApplied_GRAMD1B_padj_0_1.tsv", col.names = TRUE, row.names = FALSE, quote = FALSE, sep = '\t')
```

```{r}
head(deResultsUpdated)
```

#Create a ranked list for GSEA
```{r}
#Create ranked list for running Pre-ranked GSEA tool
rankedList = na.omit(data.frame(deResultsUpdated$Name, deResultsUpdated$log2FoldChange_lfcShrinkApplied))
rankedList = rankedList[with(rankedList, order(deResultsUpdated.log2FoldChange_lfcShrinkApplied)), ]
#Save the ranked list to a new file for GSEA.
write.table(rankedList, file = "../results/de_gsea/rankedList_GRAMD1B.rnk", col.names = FALSE, row.names = FALSE, quote = FALSE, sep = '\t')
```

#Plot survival curves between the continuous signature and the two groups
## First add necesssary columns
```{r}
suppressPackageStartupMessages(library("survival"))
# survival outcome 1: "d" for death and "ltrc" for left truncated and right censored
clinicalData$survival_d_ltrc = Surv(time  = as.numeric(difftime(clinicalData$DOE, clinicalData$DDiag))/365.25,
                                    time2 = as.numeric(difftime(clinicalData$DOC, clinicalData$DDiag))/365.25,
                                    event = ifelse(clinicalData$Dead == FALSE, 0, 1))

# survival outcome 2: "rd" for relapse or death, "ltrc" for left truncated and right censored
clinicalData$survival_rd_ltrc = Surv(time  = as.numeric(difftime(clinicalData$DOE, clinicalData$DDiag))/365.25,
                                     time2 = as.numeric(difftime(apply(clinicalData[, c("DOC", "DDistMets")],1,min,na.rm=TRUE), clinicalData$DDiag))/365.25,
                                     event = ifelse((clinicalData$Dead == FALSE & clinicalData$EventMet == "No"), 0, 1))

if(tissue=="Skin"){
  jt_ns_sm_scores = read.table("../JT_NS_SM_Combined.xls - JT_NS_SM.tsv", sep = "\t", header = TRUE, quote = "")
  combinedScore <- data.frame("Scanned_file.me"= jt_ns_sm_scores$JT_Scanned_file.me, 
                              "R.Seq_sampleID"= jt_ns_sm_scores$JT_R.Seq_sampleID, 
                              "ClarkScore"=ifelse(jt_ns_sm_scores$JT_Clark_score==jt_ns_sm_scores$NS_Clark_score,
                                                  jt_ns_sm_scores$JT_Clark_score, 
                                                  jt_ns_sm_scores$SM_Clark_score),
                           "MIAScore"= ifelse(jt_ns_sm_scores$JT_TIL_grade==jt_ns_sm_scores$NS_TIL.GRADE,
                                              jt_ns_sm_scores$JT_TIL_grade, 
                                              jt_ns_sm_scores$SM_TIL_grade))
  #Remove duplicated entries
  combinedScore<-combinedScore[!duplicated(combinedScore$R.Seq_sampleID), ]
  rownames(combinedScore)<-combinedScore$R.Seq_sampleID #rename rownames with sample ids
  combinedScore$ClarkScore <- as.factor(combinedScore$ClarkScore)
  levels(combinedScore$ClarkScore) <- c("absent", "nonbrisk", "brisk")
  clinicalData<-merge(clinicalData, combinedScore, by.x="RNA.Seq.Sample", by.y="R.Seq_sampleID")
  rownames(clinicalData)<-clinicalData$RNA.Seq.Sample
}
```

## then calculate the values
```{r}
os<-data.frame()
pfs<-data.frame()
temp<-data.frame()

combinations<-c("Signature", "SignatureGroupMean", "SignatureGroupMedian", "SignatureGroup0_33", "SignatureGroup0_66", "SignatureGroupDensityIntersection")

for (combination in combinations) {
  if (tissue=="Skin") {
    os_formula<-paste0("survival_d_ltrc~",combination,"+Sex+Age+as.numeric(Nclass)+as.character(Stage)+ECOG+treatment+ClarkScore")
    pfs_formula<-paste0("survival_rd_ltrc~",combination,"+Sex+Age+as.numeric(Nclass)+as.character(Stage)+ECOG+treatment+ClarkScore")
  } else{
    os_formula<-paste0("survival_d_ltrc~",combination,"+Sex+Age+as.numeric(Nclass)+as.character(Stage)+ECOG+treatment")
    pfs_formula<-paste0("survival_rd_ltrc~",combination,"+Sex+Age+as.numeric(Nclass)+as.character(Stage)+ECOG+treatment")
  }
  
  fit = coef(summary(coxph(as.formula(os_formula), data=clinicalData)))
  mid  = fit[1,c("exp(coef)")]
  low  = exp(fit[1,c("coef")]-
              qnorm(.975)*fit[1,c("se(coef)")])                   
  high = exp(fit[1,c("coef")]+
              qnorm(.975)*fit[1,c("se(coef)")])  
  pval = fit[1,c("Pr(>|z|)")]
  temp["Signature",c("Signature", "HR","low","high","pval")] = c(rownames(fit)[1],mid,low,high,pval)
  os<-rbind(os, temp)
  
  fit = coef(summary(coxph(as.formula(pfs_formula), data=clinicalData)))
  mid  = fit[1,c("exp(coef)")]
  low  = exp(fit[1,c("coef")]-
              qnorm(.975)*fit[1,c("se(coef)")])                   
  high = exp(fit[1,c("coef")]+
              qnorm(.975)*fit[1,c("se(coef)")])  
  pval = fit[1,c("Pr(>|z|)")]
  temp["Signature",c("Signature", "HR","low","high","pval")] = c(rownames(fit)[1],mid,low,high,pval)
  pfs<-rbind(pfs, temp)
}

write.table(os, paste("../results/survival/AVAST-M_",tissue,"_os.tsv", sep=""), sep="\t", col.names = T, row.names = F)
write.table(pfs, paste("../results/survival/AVAST-M_",tissue,"_pfs.tsv", sep=""), sep="\t", col.names = T, row.names = F)
```
